---
output: pdf_document
---

"StarBug_2016_Analysis"

R Script to combine all the .txt files from the Transect Measure output for Starbug November 2016 Ningaloo Outlook Field Trip and produce a .csv output file in a format standardized for Margaret Mille. Initial data preperation steps written by Cindy Bessey, 5/5/2015, edited and added to by Joe Turner 23/11/2016.

Prepare TransectMeasure data for analysis

```{r Data set-up, message=FALSE, warning=FALSE}

# 1) Set working directory and load packages

# FOR UWA laptop setwd("C:/Users/21901029/OneDrive/Documents/PhD/Data/Image_Analysis/TransectMeasure")

setwd("C:/Users/TUR262/OneDrive/Documents/PhD/Data/Image_Analysis/TransectMeasure")

library(plyr)
library(vegan)
library(doBy)
library(rgdal)
library(clustsig)
library(gdata)
library(ggplot2)
library(scales)
library(ggrepel)
library(RColorBrewer)
library(reshape2)
library(reshape)
library(grid)
library(dplR)
library(gridExtra)
st.err <- function(x) {sd(x)/sqrt(length(x))}

#2) Make an automatic list of all files present in the directory
file_list<-list.files()

#3) Combine all the files into a list where each layer contains only the data
##This requires the first 4 lines of the Transect Measure .txt output to be ignored
dataset <- lapply (file_list, read.table, sep="\t", header=FALSE, row.names=NULL, skip=5)

#4) Combine the data into a single dataframe
alldata<-do.call(rbind.fill,dataset)

#5)Standardize the column names

FILENAME<-alldata[,1]
FRAME<-alldata[,2]
IMAGE_ROW<-alldata[,4]
IMAGE_COL<-alldata[,5]
IMAGE_DATE<-alldata[,6]
LOCATION<-alldata[,7]
SITEID<-alldata[,8]
DEPTH<-alldata[,9]
OBSERVER<-alldata[,10]
CONTINUOUS_Y_N<-alldata[,11]
BROADGROUP<-alldata[,18]
DESCRIPTION<-alldata[,19]
FAMILY<-alldata[,20]
GENUS<-alldata[,21]
SPECIES<-alldata[,22]

#6) Combine into a final dataframe for output
final.data<-data.frame(FILENAME,FRAME,IMAGE_ROW,IMAGE_COL,IMAGE_DATE,LOCATION,SITEID,DEPTH,OBSERVER,CONTINUOUS_Y_N,BROADGROUP,DESCRIPTION,FAMILY,GENUS,SPECIES)

#Output a .csv file to the directory where all the .txt files are located
# write.csv(final.data, file="StarbugImageData2016.csv")

```

Start the data manipulation, statistical analysis and plot generation.
First load in the data for the images produced in the steps above and the environmental data from the StarBug.

```{r Load Data, message=FALSE, warning=FALSE}

setwd("C:/Users/TUR262/OneDrive/Documents/PhD/Data/Image_Analysis")

imagedata <- read.csv("StarbugImageData2016.csv", header = TRUE)
metadata <- read.csv("metadata.csv", header = TRUE)
env <- read.csv("allenvdata.csv", header = TRUE)

```

Produce the species matrix style data and calculate the percentage cover (multiply by 4 as 25 points analysed per image).
Calculated for each category type.  Broad for hard coral, family for acroporids, description for non-acroporid coral growth forms.

```{r Create % cover matrices, message=FALSE, warning=FALSE}

broad<-cast(imagedata, FILENAME ~ BROADGROUP)
broadpercent<- broad*4
broadpercent[,1]<-broad[,1]
broadpercent[,2]<- NULL
write.csv(broadpercent, file = "broadpercent.csv")

family<-cast(imagedata, FILENAME ~ FAMILY)
familypercent<- family*4
familypercent[,1]<-family[,1]
familypercent[,2]<- NULL
write.csv(familypercent, file = "familypercent.csv")

descrip<-cast(imagedata, FILENAME ~ DESCRIPTION)
descrippercent<- descrip*4
descrippercent[,1]<-descrip[,1]
descrippercent[,2]<- NULL
write.csv(descrippercent, file = "descrippercent.csv")


```


Line up the image analysis data with the environmental data.
It may be worth removing the <10m metadata for purposes of making plots as data can be erratic.

```{r Merge image data with environmental data, message=FALSE, warning=FALSE}

metadata$FILENAME<-gsub( " ", "", paste(metadata$FILENAME))
mergebroad<- merge(broadpercent, metadata, by.x = "FILENAME", by.y = "FILENAME", all.x=TRUE)
write.csv(mergebroad, file = "mergebroad.csv")

mergefamily<- merge(familypercent, metadata, by.x = "FILENAME", by.y = "FILENAME", all.x=TRUE)
write.csv(mergefamily, file = "mergefamily.csv")

mergedescrip<- merge(descrippercent, metadata, by.x = "FILENAME", by.y = "FILENAME", all.x=TRUE)
write.csv(mergedescrip, file = "mergedescrip.csv")


metadata<-metadata[metadata$X2m_depth!="<10",]


```

\newpage
Calculating coral cover by Area and 2m depth categories.  This can be done for overall hard coral (using the broad habitat categories).  Graph showing light can also be added and ploted underneath main plot. Light data is smoothed using a GAM method.

```{r Total Coral Cover, message=FALSE, warning=FALSE}

#### For Broad Habitat Categories ####

colnames(mergebroad)[6]<- "Hard_Coral"
meancoral<-aggregate(Hard_Coral ~ Area + X2m_depth, data = mergebroad, FUN= "mean")
sdcoral<-aggregate(Hard_Coral ~ Area + X2m_depth, data = mergebroad, FUN= "sd")
secoral<-aggregate(Hard_Coral ~ Area + X2m_depth, data = mergebroad, st.err)
light<-aggregate(Energy1_lic ~ Area + X2m_depth, data = metadata, FUN = "mean")
meancoral[,4]<-sdcoral[,3]
meancoral[,5]<-secoral[,3]
colnames(meancoral)[2]<- "Depth"
colnames(meancoral)[3]<- "Mean_Coral"
colnames(meancoral)[4]<- "SD_Coral"
colnames(meancoral)[5]<- "SE_Coral"


g1<- ggplot(data=meancoral,aes(x=Depth, y=Mean_Coral, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_Coral-SE_Coral, ymax=Mean_Coral+SE_Coral), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Hard Coral Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 65))

g2<-ggplot(env,aes(x = Depth1_lic, y = Energy1_lic, colour = Area)) +
  geom_smooth(se= FALSE, size = 1) +
  xlab("Depth (m)") +
  ylab ("Light Energy") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), limits = c(10, 50))

grid.newpage()
grid.draw(rbind(ggplotGrob(g1), ggplotGrob(g2), size = "last"))

```


\newpage
Coral cover by Genera / growth form (as resolution commonly only allows "non-Acropora" identification).

```{r Coral cover by growth form, message=FALSE, warning=FALSE}

#### Acropora ####

meanacrop<-aggregate(Acroporidae ~ X2m_depth, data = mergefamily, FUN= "mean")
sdacrop<-aggregate(Acroporidae ~ X2m_depth, data = mergefamily, FUN= "sd")
seacrop<-aggregate(Acroporidae ~ X2m_depth, data = mergefamily, st.err)
meanacrop[,3]<-sdacrop[,2]
meanacrop[,4]<-seacrop[,2]
meanacrop[,5]<-"Acropora"
colnames(meanacrop)[1]<- "Depth"
colnames(meanacrop)[2]<- "Mean"
colnames(meanacrop)[3]<- "SD"
colnames(meanacrop)[4]<- "SE"
colnames(meanacrop)[5]<- "Type"
meanacrop$Type<- as.factor(meanacrop$Type)

#### Encrusting non-Acropora ####

colnames(mergedescrip)[29]<- "Encrusting_Coral"
meanencrust<-aggregate(Encrusting_Coral ~ X2m_depth, data = mergedescrip, FUN= "mean")
sdencrust<-aggregate(Encrusting_Coral ~ X2m_depth, data = mergedescrip, FUN= "sd")
seencrust<-aggregate(Encrusting_Coral ~ X2m_depth, data = mergedescrip, st.err)
meanencrust[,3]<-sdencrust[,2]
meanencrust[,4]<-seencrust[,2]
meanencrust[,5]<-"Encrusting non-Acropora"
colnames(meanencrust)[1]<- "Depth"
colnames(meanencrust)[2]<- "Mean"
colnames(meanencrust)[3]<- "SD"
colnames(meanencrust)[4]<- "SE"
colnames(meanencrust)[5]<- "Type"
meanencrust$Type<- as.factor(meanencrust$Type)


#### Foliose non-Acropora ####

colnames(mergedescrip)[31]<- "Foliose_Coral"
meanfoli<-aggregate(Foliose_Coral ~ X2m_depth, data = mergedescrip, FUN= "mean")
sdfoli<-aggregate(Foliose_Coral ~ X2m_depth, data = mergedescrip, FUN= "sd")
sefoli<-aggregate(Foliose_Coral ~ X2m_depth, data = mergedescrip, st.err)
meanfoli[,3]<-sdfoli[,2]
meanfoli[,4]<-sefoli[,2]
meanfoli[,5]<-"Foliose non-Acropora"
colnames(meanfoli)[1]<- "Depth"
colnames(meanfoli)[2]<- "Mean"
colnames(meanfoli)[3]<- "SD"
colnames(meanfoli)[4]<- "SE"
colnames(meanfoli)[5]<- "Type"
meanfoli$Type<- as.factor(meanfoli$Type)


#### Massive non-Acropora ####

colnames(mergedescrip)[34]<- "Massive_Coral"
meanmass<-aggregate(Massive_Coral ~ X2m_depth, data = mergedescrip, FUN= "mean")
sdmass<-aggregate(Massive_Coral ~ X2m_depth, data = mergedescrip, FUN= "sd")
semass<-aggregate(Massive_Coral ~ X2m_depth, data = mergedescrip, st.err)
meanmass[,3]<-sdmass[,2]
meanmass[,4]<-semass[,2]
meanmass[,5]<-"Massive non-Acropora"
colnames(meanmass)[1]<- "Depth"
colnames(meanmass)[2]<- "Mean"
colnames(meanmass)[3]<- "SD"
colnames(meanmass)[4]<- "SE"
colnames(meanmass)[5]<- "Type"
meanmass$Type<- as.factor(meanmass$Type)


corals<-rbind(meanmass, meanfoli, meanacrop, meanencrust)
corals[is.na(corals)]<- 0


ggplot(data=corals,aes(x=Depth, y=Mean, colour=Type)) +
  geom_line(aes(group = Type), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth") +
  ylab("Mean Coral Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20))


```

\newpage

It is also possible to investigate sponge and macroalgal abundance with depths.

```{r Sponge and Macroalgae cover, message=FALSE, warning=FALSE}

mergebroad$Sponge <- as.numeric(mergebroad$Sponge)
mergebroad$Sponges <- as.numeric(mergebroad$Sponges)
mergebroad$AllSponge <- mergebroad$Sponge + mergebroad$Sponges

meansponge<-aggregate(AllSponge ~ Area + X2m_depth, data = mergebroad, FUN= "mean")
sdsponge<-aggregate(AllSponge ~ Area + X2m_depth, data = mergebroad, FUN= "sd")
sesponge<-aggregate(AllSponge ~ Area + X2m_depth, data = mergebroad, st.err)
meansponge[,4]<-sdsponge[,3]
meansponge[,5]<-sesponge[,3]
colnames(meansponge)[2]<- "Depth"
colnames(meansponge)[3]<- "Mean"
colnames(meansponge)[4]<- "SD"
colnames(meansponge)[5]<- "SE"


ggplot(data=meansponge,aes(x=Depth, y=Mean, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Sponge Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 15))


# Brown Macroalgae

colnames(mergefamily)[7]<- "Brown_Macroalgae"
meanbmacro<-aggregate(Brown_Macroalgae ~ Area + X2m_depth, data = mergefamily, FUN= "mean")
sdbmacro<-aggregate(Brown_Macroalgae ~ Area + X2m_depth, data = mergefamily, FUN= "sd")
sebmacro<-aggregate(Brown_Macroalgae ~ Area + X2m_depth, data = mergefamily, st.err)
meanbmacro[,4]<-sdbmacro[,3]
meanbmacro[,5]<-sebmacro[,3]
colnames(meanbmacro)[2]<- "Depth"
colnames(meanbmacro)[3]<- "Mean"
colnames(meanbmacro)[4]<- "SD"
colnames(meanbmacro)[5]<- "SE"


ggplot(data=meanbmacro,aes(x=Depth, y=Mean, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Brown Macroalgal Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40))


# Red Macroalgae

colnames(mergefamily)[26]<- "Red_Macroalgae"
meanrmacro<-aggregate(Red_Macroalgae ~ Area + X2m_depth, data = mergefamily, FUN= "mean")
sdrmacro<-aggregate(Red_Macroalgae ~ Area + X2m_depth, data = mergefamily, FUN= "sd")
sermacro<-aggregate(Red_Macroalgae ~ Area + X2m_depth, data = mergefamily, st.err)
meanrmacro[,4]<-sdrmacro[,3]
meanrmacro[,5]<-sermacro[,3]
colnames(meanrmacro)[2]<- "Depth"
colnames(meanrmacro)[3]<- "Mean"
colnames(meanrmacro)[4]<- "SD"
colnames(meanrmacro)[5]<- "SE"


ggplot(data=meanrmacro,aes(x=Depth, y=Mean, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Red Macroalgal Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 25))


```

\newpage
It may be beneficial to remove any images of substrate that is unsuitable for coral.  A threshold amount of Sand can be defined and images removed accordingly. Data can then be analysed as above to produce plots of average coral cover.

```{r, warning = FALSE, message = FALSE}

mergebroadnosand <- mergebroad[mergebroad$Sand<=24,]

# This code will overwrite the mean, sd, and se coral files

meancoral<-aggregate(Hard_Coral ~ Area + X2m_depth, data = mergebroadnosand, FUN= "mean")
sdcoral<-aggregate(Hard_Coral ~ Area + X2m_depth, data = mergebroadnosand, FUN= "sd")
secoral<-aggregate(Hard_Coral ~ Area + X2m_depth, data = mergebroadnosand, st.err)
meancoral[,4]<-sdcoral[,3]
meancoral[,5]<-secoral[,3]
colnames(meancoral)[2]<- "Depth"
colnames(meancoral)[3]<- "Mean_Coral"
colnames(meancoral)[4]<- "SD_Coral"
colnames(meancoral)[5]<- "SE_Coral"


ggplot(data=meancoral,aes(x=Depth, y=Mean_Coral, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_Coral-SE_Coral, ymax=Mean_Coral+SE_Coral), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Hard Coral Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 65))


```


\newpage
Environmental data plots including light, temperature and oxygen saturation.

```{r, warning = FALSE}

lightplot<- ggplot(env,aes(x = Depth1_lic, y = Energy1_lic, colour = Area)) +
  geom_smooth(se= FALSE, size = 1) +
  xlab("Depth (m)") +
  ylab ("Light Energy (E m-2 s-1)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0,50)) +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())


tempplot<- ggplot(env,aes(x = Depth1_ctd, y = Temp_ctd, colour = Area)) +
  geom_smooth(se= FALSE, size = 1) +
  xlab("Depth (m)") +
  ylab ("Temperature (°C)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(23, 24.5)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0,50)) +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank()) 


oxyplot<- ggplot(env,aes(x = Depth_opt , y = Saturation_opt, colour = Area)) +
  geom_smooth(se= FALSE, size = 1) +
  xlab("Depth (m)") +
  ylab ("Oxygen Saturation (%)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits=c(75,95)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0,50))  


grid.arrange(lightplot, tempplot, oxyplot, nrow = 3)

```


\newpage

Investigating differences between 5m depth bins instead of 2m.

```{r 5m Analysis, warning = FALSE, message = FALSE}

# Total Hard Coral Cover

meancoral<-aggregate(Hard_Coral ~ Area + X5m_depth, data = mergebroad, FUN= "mean")
sdcoral<-aggregate(Hard_Coral ~ Area + X5m_depth, data = mergebroad, FUN= "sd")
secoral<-aggregate(Hard_Coral ~ Area + X5m_depth, data = mergebroad, st.err)
light<-aggregate(Energy1_lic ~ Area + X5m_depth, data = metadata, FUN = "mean")
meancoral[,4]<-sdcoral[,3]
meancoral[,5]<-secoral[,3]
colnames(meancoral)[2]<- "Depth"
colnames(meancoral)[3]<- "Mean_Coral"
colnames(meancoral)[4]<- "SD_Coral"
colnames(meancoral)[5]<- "SE_Coral"


ggplot(data=meancoral,aes(x=Depth, y=Mean_Coral, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean_Coral-SE_Coral, ymax=Mean_Coral+SE_Coral), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Hard Coral Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 65))

#### Acropora ####

meanacrop<-aggregate(Acroporidae ~ X5m_depth, data = mergefamily, FUN= "mean")
sdacrop<-aggregate(Acroporidae ~ X5m_depth, data = mergefamily, FUN= "sd")
seacrop<-aggregate(Acroporidae ~ X5m_depth, data = mergefamily, st.err)
meanacrop[,3]<-sdacrop[,2]
meanacrop[,4]<-seacrop[,2]
meanacrop[,5]<-"Acropora"
colnames(meanacrop)[1]<- "Depth"
colnames(meanacrop)[2]<- "Mean"
colnames(meanacrop)[3]<- "SD"
colnames(meanacrop)[4]<- "SE"
colnames(meanacrop)[5]<- "Type"
meanacrop$Type<- as.factor(meanacrop$Type)

#### Encrusting non-Acropora ####

colnames(mergedescrip)[29]<- "Encrusting_Coral"
meanencrust<-aggregate(Encrusting_Coral ~ X5m_depth, data = mergedescrip, FUN= "mean")
sdencrust<-aggregate(Encrusting_Coral ~ X5m_depth, data = mergedescrip, FUN= "sd")
seencrust<-aggregate(Encrusting_Coral ~ X5m_depth, data = mergedescrip, st.err)
meanencrust[,3]<-sdencrust[,2]
meanencrust[,4]<-seencrust[,2]
meanencrust[,5]<-"Encrusting non-Acropora"
colnames(meanencrust)[1]<- "Depth"
colnames(meanencrust)[2]<- "Mean"
colnames(meanencrust)[3]<- "SD"
colnames(meanencrust)[4]<- "SE"
colnames(meanencrust)[5]<- "Type"
meanencrust$Type<- as.factor(meanencrust$Type)


#### Foliose non-Acropora ####

colnames(mergedescrip)[31]<- "Foliose_Coral"
meanfoli<-aggregate(Foliose_Coral ~ X5m_depth, data = mergedescrip, FUN= "mean")
sdfoli<-aggregate(Foliose_Coral ~ X5m_depth, data = mergedescrip, FUN= "sd")
sefoli<-aggregate(Foliose_Coral ~ X5m_depth, data = mergedescrip, st.err)
meanfoli[,3]<-sdfoli[,2]
meanfoli[,4]<-sefoli[,2]
meanfoli[,5]<-"Foliose non-Acropora"
colnames(meanfoli)[1]<- "Depth"
colnames(meanfoli)[2]<- "Mean"
colnames(meanfoli)[3]<- "SD"
colnames(meanfoli)[4]<- "SE"
colnames(meanfoli)[5]<- "Type"
meanfoli$Type<- as.factor(meanfoli$Type)


#### Massive non-Acropora ####

colnames(mergedescrip)[34]<- "Massive_Coral"
meanmass<-aggregate(Massive_Coral ~ X5m_depth, data = mergedescrip, FUN= "mean")
sdmass<-aggregate(Massive_Coral ~ X5m_depth, data = mergedescrip, FUN= "sd")
semass<-aggregate(Massive_Coral ~ X5m_depth, data = mergedescrip, st.err)
meanmass[,3]<-sdmass[,2]
meanmass[,4]<-semass[,2]
meanmass[,5]<-"Massive non-Acropora"
colnames(meanmass)[1]<- "Depth"
colnames(meanmass)[2]<- "Mean"
colnames(meanmass)[3]<- "SD"
colnames(meanmass)[4]<- "SE"
colnames(meanmass)[5]<- "Type"
meanmass$Type<- as.factor(meanmass$Type)


corals<-rbind(meanmass, meanfoli, meanacrop, meanencrust)
corals[is.na(corals)]<- 0


ggplot(data=corals,aes(x=Depth, y=Mean, colour=Type)) +
  geom_line(aes(group = Type), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth") +
  ylab("Mean Coral Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 20))

# Sponges

meansponge<-aggregate(AllSponge ~ Area + X5m_depth, data = mergebroad, FUN= "mean")
sdsponge<-aggregate(AllSponge ~ Area + X5m_depth, data = mergebroad, FUN= "sd")
sesponge<-aggregate(AllSponge ~ Area + X5m_depth, data = mergebroad, st.err)
meansponge[,4]<-sdsponge[,3]
meansponge[,5]<-sesponge[,3]
colnames(meansponge)[2]<- "Depth"
colnames(meansponge)[3]<- "Mean"
colnames(meansponge)[4]<- "SD"
colnames(meansponge)[5]<- "SE"


ggplot(data=meansponge,aes(x=Depth, y=Mean, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Sponge Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 15))


# Brown Macroalgae

colnames(mergefamily)[7]<- "Brown_Macroalgae"
meanbmacro<-aggregate(Brown_Macroalgae ~ Area + X5m_depth, data = mergefamily, FUN= "mean")
sdbmacro<-aggregate(Brown_Macroalgae ~ Area + X5m_depth, data = mergefamily, FUN= "sd")
sebmacro<-aggregate(Brown_Macroalgae ~ Area + X5m_depth, data = mergefamily, st.err)
meanbmacro[,4]<-sdbmacro[,3]
meanbmacro[,5]<-sebmacro[,3]
colnames(meanbmacro)[2]<- "Depth"
colnames(meanbmacro)[3]<- "Mean"
colnames(meanbmacro)[4]<- "SD"
colnames(meanbmacro)[5]<- "SE"


ggplot(data=meanbmacro,aes(x=Depth, y=Mean, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Brown Macroalgal Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 40))


# Red Macroalgae

colnames(mergefamily)[26]<- "Red_Macroalgae"
meanrmacro<-aggregate(Red_Macroalgae ~ Area + X5m_depth, data = mergefamily, FUN= "mean")
sdrmacro<-aggregate(Red_Macroalgae ~ Area + X5m_depth, data = mergefamily, FUN= "sd")
sermacro<-aggregate(Red_Macroalgae ~ Area + X5m_depth, data = mergefamily, st.err)
meanrmacro[,4]<-sdrmacro[,3]
meanrmacro[,5]<-sermacro[,3]
colnames(meanrmacro)[2]<- "Depth"
colnames(meanrmacro)[3]<- "Mean"
colnames(meanrmacro)[4]<- "SD"
colnames(meanrmacro)[5]<- "SE"


ggplot(data=meanrmacro,aes(x=Depth, y=Mean, colour=Area)) +
  geom_line(aes(group = Area), size = 1.5) + 
  geom_point(size=2) +
  geom_errorbar(aes(ymin=Mean-SE, ymax=Mean+SE), width=.1) +
  xlab("Depth (m)") +
  ylab("Mean Red Macroalgal Cover (% ± S.E)") +
  theme_bw() +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 25))


```



\newpage

Statistical Tests. First investigate the data

```{r}

# mergebroad$Hard_Coral <- as.numeric(mergebroad$Hard_Coral)
# ggplot(data=mergebroad, aes(mergebroad$Hard_Coral)) + 
  # geom_histogram()
# Obvious Poisson distribution

# anova(Hard_Coral ~ Area + 2m_Depth, data = mergebroad)


# BIOENV

# data<-mergebroad[c(2:6,8:13)]
# rownames(data) <- mergebroad[,1]
# envdata<-mergebroad[c(23,39,40,42,43,44,49,57,58,66,67,68)]
# rownames(envdata) <- mergebroad[,1]

# x<- bioenv(data, envdata, method = "spearman", index = "bray")
# bioenvdist(x, which = "best")



# data2<-mergefamily[c(2:14,16:33)]
# rownames(data2) <- mergefamily[,1]
# x2<- bioenv(data2, envdata, method = "spearman", index = "bray")




# Multi Dimensional Scaling




```


```{r USBL data}

library(lubridate)
library(zoo)

setwd("C:/Users/TUR262/OneDrive/Documents/PhD/PhD_Chapters/Chapter_2_Mesophotic_Communities/Data/USBL")

file_list<-list.files(pattern="*.txt")

    DF <- NULL
        for (f in file_list) {
        dat <- read.table(f, header=T, sep=",", colClasses = "character")
        dat$site<- rep(paste(f), nrow(dat))
        DF <- rbind(DF, dat)
        }
    
unique<- subset(DF, !duplicated(Easting))

write.csv(unique, file = "USBL_Unique.csv")
write.csv(DF, file = "USBL_All_Sites.csv")
write.table(unique, file = "USBL_Unique.txt", sep= ",", row.names=FALSE)

unique$date <- ifelse(unique$site== "hb01_fs.txt", "20161102",
        ifelse(unique$site== "hb01_fs.txt", "20161102",
        ifelse(unique$site== "hb02_2nd.txt", "20161102",
        ifelse(unique$site== "hb03.txt", "20161102",
        ifelse(unique$site== "hb04.txt", "20161102",
        ifelse(unique$site== "mg01.txt", "20161104",
        ifelse(unique$site== "mg02.txt", "20161104",
        ifelse(unique$site== "mg03.txt", "20161104",
        ifelse(unique$site== "tb01.txt" & unique$Msg.Time > 42677.33329, "20161103",
        ifelse(unique$site== "tb01.txt" & unique$Msg.Time < 42677.33329, "20161102",      
        ifelse(unique$site== "tb02.txt", "20161103",
        ifelse(unique$site== "tb03.txt", "20161103",
        ifelse(unique$site== "mu01.txt", "20161104",
        ifelse(unique$site== "mu02.txt", "20161104",
        ifelse(unique$site== "mu03.txt", "20161104",
                        NA  ))))))))))))))) 

unique$Time <- as.POSIXct(paste(unique$date, unique$UTC.Time), format="%Y%m%d %H%M%S", tz = "UTC")
unique$Local_Time<- unique$Time + hours(9)

write.csv(unique, file = "USBL_Unique.csv")

setwd("C:/Users/TUR262/OneDrive/Documents/PhD/PhD_Chapters/Chapter_2_Mesophotic_Communities/Data")
images<-read.csv("All_Images_and_Env_Merge.csv", header = T)
images$Local_Time <- as.POSIXct(images$Date_Time.1, format="%Y%m%d %H%M%S", tz = "Asia/Tokyo") # 9 hour difference
images$UTC.Time <- with_tz(images$Local_Time, tz = "UTC")

write.csv(images, file = "images.csv")

merge<-merge(unique, images, by.x = "Time", by.y = "UTC.Time")
merge[c(2:3, 7:8, 10)] <- list(NULL)
write.csv(merge, file = "merge.csv")


# Interpolate http://stackoverflow.com/questions/26806140/interpolate-coordinates-at-unrecorded-timestamps
# Need to do for each site seperately
# Below is linear

hb01_fs<- subset(unique, site == "hb01_fs.txt")
full.time    <- with(hb01_fs,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(hb01_fs[,3:4],hb01_fs$Time)        # convert to zoo object
resulthb01 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulthb01)

hb02_2nd<- subset(unique, site == "hb02_2nd.txt")
full.time    <- with(hb02_2nd,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(hb02_2nd[,3:4],hb02_2nd$Time)        # convert to zoo object
resulthb02 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulthb02)

hb03<- subset(unique, site == "hb03.txt")
full.time    <- with(hb03,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(hb03[,3:4],hb03$Time)        # convert to zoo object
resulthb03 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulthb03)

hb04<- subset(unique, site == "hb04.txt")
full.time    <- with(hb04,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(hb04[,3:4],hb04$Time)        # convert to zoo object
resulthb04 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulthb04)

mg01<- subset(unique, site == "mg01.txt")
full.time    <- with(mg01,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(mg01[,3:4],mg01$Time)        # convert to zoo object
resultmg01 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resultmg01)

mg02<- subset(unique, site == "mg02.txt")
full.time    <- with(mg02,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(mg02[,3:4],mg02$Time)        # convert to zoo object
resultmg02 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resultmg02)

mg03<- subset(unique, site == "mg03.txt")
full.time    <- with(mg03,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(mg03[,3:4],mg03$Time)        # convert to zoo object
resultmg03 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resultmg03)

mu01<- subset(unique, site == "mu01.txt")
full.time    <- with(mu01,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(mu01[,3:4],mu01$Time)        # convert to zoo object
resultmu01 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resultmu01)

mu02<- subset(unique, site == "mu02.txt")
full.time    <- with(mu02,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(mu02[,3:4],mu02$Time)        # convert to zoo object
resultmu02 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resultmu02)

mu03<- subset(unique, site == "mu03.txt")
full.time    <- with(mu03,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(mu03[,3:4],mu03$Time)        # convert to zoo object
resultmu03 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resultmu03)

tb01<- subset(unique, site == "tb01.txt")
full.time    <- with(tb01,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(tb01[,3:4],tb01$Time)        # convert to zoo object
resulttb01 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulttb01)

tb02<- subset(unique, site == "tb02.txt")
full.time    <- with(tb02,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(tb02[,3:4],tb02$Time)        # convert to zoo object
resulttb02 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulttb02)

tb03<- subset(unique, site == "tb03.txt")
full.time    <- with(tb03,seq(Time[1],tail(Time,1),by=1))
df.zoo <- zoo(tb03[,3:4],tb03$Time)        # convert to zoo object
resulttb03 <- na.approx(df.zoo,xout=full.time)  # interpolate; result is also a zoo object
head(resulttb03)


setwd("C:/Users/TUR262/OneDrive/Documents/PhD/PhD_Chapters/Chapter_2_Mesophotic_Communities/Data/USBL/Interpolated")
write.zoo(resulthb01, file = "hb01_fs_Interpolation.txt")
write.zoo(resulthb02, file = "hb02_2nd_Interpolation.txt")
write.zoo(resulthb03, file = "hb03_Interpolation.txt")
write.zoo(resulthb04, file = "hb04_Interpolation.txt")
write.zoo(resultmg01, file = "mg01_Interpolation.txt")
write.zoo(resultmg02, file = "mg02_Interpolation.txt")
write.zoo(resultmg03, file = "mg03_Interpolation.txt")
write.zoo(resultmu01, file = "mu01_Interpolation.txt")
write.zoo(resultmu02, file = "mu02_Interpolation.txt")
write.zoo(resultmu03, file = "mu03_Interpolation.txt")
write.zoo(resulttb01, file = "tb01_Interpolation.txt")
write.zoo(resulttb02, file = "tb02_Interpolation.txt")
write.zoo(resulttb03, file = "tb03_Interpolation.txt")

temp <- list.files(pattern="*.txt")
    Interp <- NULL
        for (f in temp) {
        dat <- read.csv(f, header=F, sep=" ", colClasses = "character")
        dat$site<- rep(paste(f), nrow(dat))
        Interp <- rbind(Interp, dat)
        }
    
colnames(Interp)<-c("Date", "Time", "Easting", "Northing", "Site")
Interp<-Interp[ which( ! Interp$Date %in% "Index") , ]
Interp$Date<- as.factor(Interp$Date)
Interp$Time<- as.factor(Interp$Time)
Interp$UTC_Time <- as.POSIXct(paste(Interp$Date, Interp$Time), format="%Y-%m-%d %H:%M:%S", "UTC")
write.csv(Interp, "Interp_All_Sites.csv")

merge<-merge(Interp, images, by.x = "UTC_Time", by.y = "UTC.Time")
merge[c(2:3, 6:10)] <- list(NULL)

setwd("C:/Users/TUR262/OneDrive/Documents/PhD/PhD_Chapters/Chapter_2_Mesophotic_Communities/Data")
write.csv(merge, file = "merge.csv")
helbybanks<- subset(merge, Area == "Helby Banks")
othersites<- subset(merge, Area != "Helby Banks")
write.csv(helbybanks, file = "Helby_Banks_Sites_Image_Merge_UTM50S.csv")
write.csv(othersites, file = "Other_Sites_Image_Merge_UTM49S.csv")


# Polynomial interpolation of coordinates in r 
splinefun()

```

